Convective dynamics with mixed temperature boundary conditions: \\ why thermal relaxation matters and how to accelerate it

### 2020-05-20 tex/mixedft_evolution_ms.tex
lots of small tweaks to the text

\begin{abstract}
Astrophysical simulations of convection frequently impose different thermal boundary conditions at the top and the bottom of the domain in an effort to more accurately reflect the natural system being modeled.
In this work, we study \RB convection (RBC) \ea{under the Boussinesq approximation}.
\ea{We examine simulations with mixed temperature boundary conditions in which the flux is fixed at the bottom boundary and the temperature is fixed at the top (``FT'').}
We aim to understand how FT boundaries change the nature of the convective solution compared to the traditional choice of \ea{thermal boundaries, in which the temperature is fixed} at the top and bottom of the domain (``TT'').
\ea{We demonstrate that the timescale of thermal relaxation for FT simulations is dependent upon the initial conditions.
``Classic'' initial conditions which employ a hydrostatically- and thermally- balanced linear temperature profile exhibit a long thermal relaxation.
This long relaxation timescale can be skipped by either using a TT simulation's nonlinear state as initial conditions for an FT simulation, or by choosing initial conditions which more accurately guess the temperature difference between the top and bottom plates in the evolved state.}
In the \ea{thermally} relaxed\ea{, statistically stationary} state, the mean behavior of an FT simulation corresponds to an equivalent simulation with TT boundaries\ea{, and time- and volume-averaged flow statistics like the Nusselt number and the P\'{e}clet number are indistinguishable between FT and TT simulations}.
\ea{FT boundaries are fundamentally asymmetric, and we examine the asymmetries that these boundaries produce in the flow.
We find that the fixed-flux boundary produces more extreme temperature events than the fixed-temperature boundary.
However, these near-boundary asymmetries do not measurably break the symmetry in the convective interior.
We briefly explore rotating RBC to demonstrate that our findings with respect to thermal relaxation carry over to this more complex case.}
\end{abstract}
\maketitle

@@

\ea{
\subsection{Timescales}
\label{sec:timescales}
}
\subsubsection{Dynamical Timescales}
One \ea{result of the mapping in Eqn.~\ref{eqn:ra_relation}} is that the nondimensional dynamical freefall timescale is a poor description of the \emph{evolved} freefall timescale.
The velocities in an FT simulation are smaller than the velocities in a TT simulation by a factor of $\sqrt{\text{Nu}}$.
As a result, every nondimensional simulation freefall time unit in a TT simulation samples a factor of $\sqrt{\text{Nu}}$ more dynamics than a time unit in an FT simulation.
\ea{We therefore define the \emph{evolved} freefall time, $\tau_{\text{ff, ev}} = \sqrt{\text{Nu}}$ (for FT simulations) or $\tau_{\text{ff, ev}} = 1$ (for TT simulations).
To ensure accurate comparisons, we will measure flow statistics over multiples of $\tau_{\text{ff, ev}}$ rather than over multiples of the nondimensional time units.}

\ea{
\subsubsection{Evolutionary Timescales}
The thermal energy reservoir of our 3D Cartesian systems, and the rate of change of this reservoir due to conduction at the boundaries, is
\begin{equation}
E = \iiint T\,dx\,dy\,dz = \Gamma^2 \int T\,dz,
\qquad
\frac{dE}{dt} = -\iiint \grad\cdot F_{\text{cond}}\,dx\,dy\,dz = -\Gamma^2 F_{\text{cond,z}}\bigg|_{z=-0.5}^{z=0.5}
\end{equation}
where $F_{\text{cond}} = -\Peff^{-1}\grad T$ is the conductive flux whose z-component is $F_{\text{cond,z}}$, and where $\int T dz = \Delta T / 2$ in RBC.
The thermal relaxation time of an RBC experiment is therefore
\begin{equation}
\tau_{\text{th}} = \frac{\Delta E}{dE/dt},
\,\,\text{with}\,\,
\Delta E = E(t=\infty) - E(t=0).
\end{equation}
In the case of TT boundary conditions, we expect $\Delta E = 0$, as the initial and final state have the same $\Delta T$.
The goal of our ``Nu-based'' and TT-to-FT initial conditions is to create FT systems in which $\Delta E \approx 0$, creating a system with a negligible relaxational timescale.

For the case of classic ICs in an FT system, $\Delta T(t=0) = -1$ and $\Delta T(t=\infty) = -\text{Nu}^{-1}$, and the temperature gradient at the bottom boundary is fixed at a value of -1.
The thermal relaxation timescale is therefore
\begin{equation}
\tau_{\text{th, FT-classic}} = \Peff \frac{\text{Nu}^{-1} - 1}{\partial_z T(z=0.5) + 1} \sim \frac{\sqrt{\text{Ra Pr}}}{|\partial_z T(z = 0.5)| - 1},
\end{equation}
where the final expression is for the large Ra case where Nu$^{-1} \ll 1$.
The magnitude of the temperature derivative achieved at the top boundary is initially very large ($|\partial_z T(z=0.5)|_{\text{early}} > \text{Nu}(t=\infty)$; see e.g., Fig.~1b in \cite{anders&all2018}), but decreases in magnitude throughout the evolution of a simulation, making it difficult to estimate the true thermal relaxation time.
However, it is reasonable to assume that the evolutionary timescale lies within the window $\sqrt{\text{Ra Pr}} \,\text{Nu}^{-1} \lesssim \tau_{\text{th, FT-classic}} \lesssim \sqrt{\text{Ra Pr}}$.
In practice, we find that our non-rotating simulations equilibrate in $\sim 2.5\sqrt{\text{Ra}\,\text{Pr}}/\text{Nu}$ nondimensional time units (see section \ref{sec:time_evolution_classic_ICs}), and our rotating simulation equilibrates in $\sim (2/3)\sqrt{\text{Ra}\,\text{Pr}}/\text{Nu}$ (see section \ref{sec:results_rotating}).

@@

\subsection{TT-to-FT \& Nu-based ICs: rapidly equilibrated FT simulations}
\label{sec:tt-to-ft}

\begin{figure}
\includegraphics[width=\textwidth]{./figs/rbc_restart_description.pdf}
\caption{ 
	(Left three panels) Time traces of scalar quantities, which have been averaged over a rolling time window of 25 simulation time units, are shown for a simulation with Ra$_{\Delta T} = 10^9$ which starts with TT boundary conditions and then is switched to FT boundary conditions \ea{with Ra$_{\partial_z T} = 4.83 \times 10^{10}$}.
	The time of the change of boundary conditions is denoted by the vertical black line.
	(Top panel) The evolution of Ra is shown; Ra$_{\partial_z T}/(4.83\times 10^{10})$ is shown as a dashed-dot line, while Ra$_{\Delta T}/10^9$ is shown as a solid line.
	The mean value of temperature gradient (middle panel) and temperature (bottom panel) at the bottom boundary are also shown.
	In the TT initial state, the temperature is held constant at a value of 1 and the temperature derivative fluctuates around a value of $\text{Nu}$.
	In the FT final state, the temperature derivative is held constant at a value of -1 and the temperature value fluctuates around a value of $\text{Nu}^{-1}$.
	(right four panels) PDFs are shown which compare TT-to-FT dynamics (black PDFs) to dynamics from \ea{the classic-IC} FT case from Fig.~\ref{fig:rbc_scalar_comparisons} (green PDFs).
	We display the temperature field (upper left), enstrophy (upper right), nonlinear convective enthalpy flux (bottom left), and vertical velocity (bottom right).
\label{fig:rbc_restart_description} }
\vspace{-0.25cm}
\end{figure}

@@

\ea{
\subsubsection{Nu-based ICs}
The time evolution of FT simulations with Nu-based ICs is similar to the time evolution of TT simulations with classic ICs, with a few caveats.
The interior temperature gradient ($\grad_{\text{interior}}$ in Eqn.~\ref{eqn:nu_based_gradT}) is slightly positive at high Ra, which means that the interior is marginally thermally stable while the boundary layers are thermally unstable.
During the convective transient, plumes from the boundaries eat away at this interior stratification over a few tens of freefall timescales, after which time the interior is well mixed and a classic roll solution is achieved.
This behavior is quite different from classic ICs in TT simulations, in which the full domain is initially unstable and a roll solution is obtained immediately after the onset of nonlinear convection.
Regardless, the time required for the temperature field to reach statistical equilibrium is a few tens of freefall times rather than a few thousands of freefall times for classic ICs.
Per table \ref{table:pdf_values}, the evolved state of FT simulations with Nu-based ICs and classic ICs are very similar.
}

\subsection{\ea{Discussion of 2D results}} 
We note briefly that \ea{Nu-based and} TT-to-FT \ea{ICs are} only \ea{two} of many ways of accelerating the thermal relaxation of an FT simulation.
We discuss other mechanisms, and explore one in detail, in our previous work \cite{anders&all2018}.
We note however that the TT-to-FT setup described here is likely the least complicated mechanism for achieving rapid relaxation in a simplified RBC setup that we are aware of.
The successful degree with which this mechanism reproduces the evolved dynamics suggests that thermal relaxation occurs in two parts:
\begin{enumerate}
\item Changes to the simulation energy reservoir, and
\item Restratification of the experiment.
\end{enumerate}
The thermal energy reservoir of TT simulations does not change between the initial and final state.
The rapid relaxation of TT simulations\ea{, as well as FT simulations with Nu-based and TT-to-FT ICs}, therefore suggests that experimental restratification occurs rapidly in RBC. 
The long rundown of \ea{classic} FT experiments on display in Fig.~\ref{fig:rbc_scalar_comparisons} is entirely due to the energy reservoir (the temperature jump across the domain) drifting over time.

\ea{
As a final note, we find that measures of the velocity field (e.g., the kinetic energy and Pe) take a few hundred freefall timescales to relax to their final value in TT and FT (Nu-based IC) simulations in 2D at high values of Ra$_{\Delta T}\,\gtrsim 10^{8.67}$.
This velocity field relaxation happens despite instant \emph{thermal} equilibration of the temperature field for these simulations.
We find that the kinetic energy increases by less than a factor of two from its initial post-transient value to its final value in the equilibrated state.
}

@@

\ea{
\subsection{3D verification of 2D results}
Timestepping through the thermal relaxation of a classic FT simulation is prohibatively expensive in 3D, even for our least turbulent case with an evolved Ra$_{\Delta T} \approx 10^8$.
In order to timestep through the intense, early convective transient, we anticipate that such a simulation would require a coefficient resolution of $\gtrsim 512^3$, a factor of 64 times more coefficients than our comparable 2D case at 2048x1024.
This 2D, classic-IC, FT case took $1.21 \times 10^5$ cpu-hours, so we estimate that a classic FT case in 3D at this value of Ra$_{\partial_z T}$ would take O(10 million) cpu-hours.
We found in previous work \cite{anders&all2018} at lower Ra$_{\partial_z T}$ that 3D classic FT cases exhibited the same thermal rundown as 2D cases.
As a result, we will focus here on 3D comparisons of FT and TT simulations in the statistically-stationary state through the use of TT-to-FT initial conditions.

We find that equilibrated FT and TT simulations are analogous in a volume-averaged sense: measurements of Nu (31.4 for TT, 31.9 for FT) and Pe ($1.78 \times 10^3$ for both) are nearly indistinguishable.
In Fig.~\ref{fig:rbc_3D_panels}, we show dynamical slices from 3D TT and TT-to-FT simulations.
As in 2D, we find no difference between the PDFs of evolved quantities like the nonlinear transport ($wT$) between TT and FT simulations.
Therefore, we have focused on the temperature fields in this figure, which does display differences between the two sets of boundary conditions.
Near the top boundary and in the interior, the TT and FT dynamics are quite similar.
However, near the bottom boundary, the warm upflows in FT simulations are hotter relative to their counterparts in TT simulations.

@@

We now extend our study to a more complicated experiment: 3D rotating RBC with an Ekman number of $10^{-6}$.
We study a TT case at Ra$_{\Delta T} = 2.75\times 10^9$, and an FT case at $\text{Ra}_{\partial_z T} = 2.1 \times 10^{10}$ (the supercriticality of the TT case is $\sim 3$).
These simulations employ stress free boundary conditions which allow for the generation of mean flows such as large scale vortices (LSV) \ea{\citep{stellmach&all2014, rubio&all2014, guervilly&all2014, guervilly&hughes2017, favier&all2014, favier&all2019, couston&all2019}}.
\ea{
Rotating convection is an excellent testbed for our TT-to-FT method.
In rotating convection, at a fixed value of Ek, as Ra increases, flows transition from the rotationally constrained to the rotationally unconstrained regime.
The scaling of Nu~vs.~Ra changes drastically between these two regimes, and is some blend of the two in the intermediary, marginally constrained regime.
Furthermore, the precise scaling law attained in the rotationally constrained regime differs as Ek changes, and these scaling laws are less straightforward and well-understood than their non-rotationally-constrained counterparts \cite{king&all2009, schmitz&tilgner2009, zhong&ahlers2010, julien&all2012, stevens&all2013, ecke2015, grooms2015, grooms&whitehead2015,  plumley&julien2019}.
As a result, the powerlaw used for Nu-based ICs would have to be a complex function of Ek, Ra, and boundary conditions for rotating RBC.
However, TT-to-FT should work generally for all parameters, so long as a TT simulation can be performed.
}

In the left three panels of Fig.~\ref{fig:rotating_panels}, we compare the time evolution of \ea{a classic-IC} FT and TT case.
The top left panel shows the evolution of Ra$_{\partial_z T}$ and Ra$_{\Delta T}$.
Even in the presence of strong rotation, the TT simulation immediately equilibrates, but the FT case takes thousands of freefall times to achieve thermal relaxation.
In the middle panel, we show the evolution of Ro; the evolved flows in both simulations exhibit rotationally constrained dynamics with Ro $\,\approx 0.1$, but the flows in the FT simulation relax to this state from an initially unconstrained state (Ro $\,\approx 1$).
This implies that the thermal relaxation process can walk through the parameter space of flow balances (e.g., the balance between Inertial and Coriolis forces) in addition to the Ra$_{\Delta T}$ parameter space.
In the bottom panel, we display the evolution of Pe over time.
Strangely, the peak value of Pe occurs a few hundred freefall times after the convective transient.
After achieving this peak value, Pe monotonically decreases toward its relaxed state.
\ea{We find that the thermal relaxation of this case takes $\sim (2/3)\sqrt{\text{Ra}_{\partial_z T}}/\text{Nu}$ nondimensional freefall time units.}

\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/rotating_panels.pdf}
\caption{ 
	(Left three panels) Time traces of scalar quantities in a \ea{classic-IC} FT (orange) and TT (purple) simulation are shown.
	All traces have been averaged over a rolling window of 50 freefall time units to increase the clarity of the evolutionary trend.
	(Top panel) Ra, normalized by the input Ra$_{\Delta T}$ = 2.75$\,\times 10^9$ of the TT simulation.
	(Middle panel) Ro evolution of both simulations; the bulk flow of the FT simulation transitions from a marginally rotationally unconstrained state to a constrained state, while the TT simulation is always rotationally constrained.
	(Bottom panel) Pe evolution of the simulations is shown, normalized by the mean value measured over the last 500 freefall times of the TT simulation.
	(Upper right panel) Parameter space of Nu vs.~Ra$_{\Delta T}$ in rotating convection.
	Circular and diamond data points are respectively simulations and experimental data points from ref.~\cite{cheng&all2015}.
	The color of the data points signifies the Ekman number of the points, and black points are non-rotating.
	Data from our Ek$\,=\,10^{-6}$ FT experiment are shown as a thick orange line with a cyan interior, where the black arrows give the direction of time.
	Some TT simulations are shown as purple circles with a cyan interior, and the TT case which corresponds to the relaxed state of the FT simulation is shown as a star.
	(Bottom right panels) Snapshots of the vertically integrated z-component of the vorticity from the FT simulation.
	At early times (left panel), a powerful large scale vortex with positive vorticity develops.
	This vortex slowly decays and becomes a vortex pair (middle panel), as seen in ref.~\cite{stellmach&all2014}.
	In the converged state, we see oscillatory behavior between this vortex pair behavior and jets (right panel).
	The TT case exhibits the oscillatory behavior between vortex pairs and jets throughout its whole evolution.
	The three vertical black lines in the left panels signify the times at which these snapshots are taken.
\label{fig:rotating_panels} }
\end{figure}

@@

\ea{
We performed a TT-to-FT simulation starting from the evolved state of the TT simulation, and its time evolution matched that seen in Fig.~\ref{fig:rbc_restart_description}.
There was no long thermal evolutionary timescale.
In the equilibrated classic FT simulation, we measured Nu$=7.90 \pm 0.01$, Pe$=1.70 \times 10^3$, and Ro$=0.115 \pm 0.006$.
In the TT-to-FT simulation, we measured Nu$=7.83 \pm 0.01$, Pe$=1.65 \times 10^3$, and Ro=$0.114 \pm 0.005$.
These Ro measurements are indistinguishable, Nu measurements are within 1\% of each other, and Pe measurements are within 3\% of each other, representing excellent agreement.
}

We once again find it important to briefly note the difference in computational cost between the \ea{classic-IC} FT, \ea{TT-to-FT,} and TT simulations conducted here.
The TT simulation shown in the left panels of Fig.~\ref{fig:rotating_panels} only cost $2.2 \times 10^3$ cpu-hours to run 5000 freefall times.
\ea{The TT-to-FT simulation cost an additional $1.28 \times 10^3$ cpu-hours to run 1000$\tau_{\text{ff, ev}}$.}
By comparison, the cost of the FT simulation shown in the same panels was \emph{roughly three orders of magnitude larger}---$2.3 \times 10^6$ cpu-hours.
The TT \ea{and TT-to-FT} simulations \ea{had a} coefficient resolution \ea{of} $128^3$.
The FT simulation's initial resolution required to resolve the convective transient was $512\times384^2$ coefficients.
We reduced the resolution to $256\times384^2$ after 100 freefall times, and then later to $128\times384^2$ after $\sim3.3 \times 10^3$ freefall times.
At each of these times, we found that lowering the \emph{horizontal} coefficient resolution of the simulation did not reproduce the simulation solution with fidelity.
This suggests that small scale turbulent velocity structures---which are injected by the vigorous transient and perhaps associated with the LSV---are long lived throughout the thermal evolution of the simulation.

>>>>

\begin{abstract}
Astrophysical simulations of convection frequently impose different thermal boundary conditions at the top and the bottom of the domain in an effort to more accurately model natural systems.
In this work, we study \RB convection (RBC) \ea{under the Boussinesq approximation}.
\ea{We examine simulations with mixed temperature boundary conditions in which the flux is fixed at the bottom boundary and the temperature is fixed at the top (``FT'').}
We aim to understand how FT boundaries change the nature of the convective solution compared to the traditional choice of \ea{thermal boundaries, in which the temperature is fixed} at the top and bottom of the domain (``TT'').
\ea{We demonstrate that the timescale of thermal relaxation for FT simulations is dependent upon the initial conditions.
``Classic'' initial conditions which employ a hydrostatically- and thermally- balanced linear temperature profile exhibit a long thermal relaxation.
This long relaxation is not seen in FT simulations which use a TT simulation's nonlinear state as initial conditions (``TT-to-FT'').}
In the \ea{thermally} relaxed\ea{, statistically stationary} state, the mean behavior of an FT simulation corresponds to an equivalent simulation with TT boundaries\ea{, and time- and volume-averaged flow statistics like the Nusselt number and the P\'{e}clet number are indistinguishable between FT and TT simulations}.
\ea{FT boundaries are fundamentally asymmetric, and we examine the asymmetries that these boundaries produce in the flow.
We find that the fixed-flux boundary produces more extreme temperature events than the fixed-temperature boundary.
However, these near-boundary asymmetries do not measurably break the symmetry in the convective interior.
We briefly explore rotating RBC to demonstrate that our findings with respect to thermal relaxation carry over to this more complex case, and to show the power of TT-to-FT initial conditions.}
\end{abstract}
\maketitle

@@

\ea{
\subsection{Timescales}
\label{sec:timescales}
}
One \ea{result of the mapping in Eqn.~\ref{eqn:ra_relation}} is that the nondimensional dynamical freefall timescale is a poor description of the \emph{evolved} freefall timescale.
The velocities in an FT simulation are smaller than the velocities in a TT simulation by a factor of $\sqrt{\text{Nu}}$.
As a result, every nondimensional simulation freefall time unit in a TT simulation samples a factor of $\sqrt{\text{Nu}}$ more dynamics than a time unit in an FT simulation.
\ea{We therefore define the \emph{evolved} freefall time, $\tau_{\text{ff, ev}} = \sqrt{\text{Nu}}$ (for FT simulations) or $\tau_{\text{ff, ev}} = 1$ (for TT simulations).
To ensure accurate comparisons, we will measure flow statistics over multiples of $\tau_{\text{ff, ev}}$ rather than over multiples of the nondimensional time units.}

\ea{
The thermal energy reservoir of our 3D Cartesian systems, and the rate of change of this reservoir due to conduction at the boundaries, is
\begin{equation}
E = \iiint T\,dx\,dy\,dz = \Gamma^2 \int T\,dz,
\qquad
\frac{dE}{dt} = -\iiint \grad\cdot \bm{F}_{\text{cond}}\,dx\,dy\,dz = -\Gamma^2 F_{\text{cond,z}}\bigg|_{z=-0.5}^{z=0.5}
\end{equation}
where $\bm{F}_{\text{cond}} = -\Peff^{-1}\grad T$ is the conductive flux whose z-component is $F_{\text{cond,z}}$, and where $\int T dz = \Delta T / 2$ in RBC.
The thermal relaxation time of an RBC experiment is therefore
\begin{equation}
\tau_{\text{th}} = \frac{\Delta E}{dE/dt},
\,\,\text{with}\,\,
\Delta E = E(t=\infty) - E(t=0).
\end{equation}
In the case of TT boundary conditions, we expect $\Delta E = 0$, as the initial and final state have the same $\Delta T$.
The goal of our ``Nu-based'' and TT-to-FT initial conditions is to create FT systems in which $\Delta E \approx 0$, creating a system with a negligible relaxational timescale.

For the case of classic ICs in an FT system, $\Delta T(t=0) = -1$ and $\Delta T(t=\infty) = -\text{Nu}^{-1}$, and the temperature gradient at the bottom boundary is fixed at a value of -1.
The thermal relaxation timescale is therefore
\begin{equation}
\tau_{\text{th, FT-classic}} = \Peff \frac{\text{Nu}^{-1} - 1}{\partial_z T(z=0.5) + 1} \sim \frac{\sqrt{\text{Ra Pr}}}{|\partial_z T(z = 0.5)| - 1},
\end{equation}
where the final expression is for the large Ra case where Nu$^{-1} \ll 1$.
The magnitude of the temperature derivative achieved at the top boundary is initially very large ($|\partial_z T(z=0.5)|_{\text{early}} > \text{Nu}(t=\infty)$; see e.g., Fig.~1b in ref.~\cite{anders&all2018}), but decreases in magnitude throughout the evolution of a simulation, making it difficult to estimate the true thermal relaxation time.
However, it is reasonable to assume that the evolutionary timescale lies within the window $\sqrt{\text{Ra Pr}} \,\text{Nu}^{-1} \lesssim \tau_{\text{th, FT-classic}} \lesssim \sqrt{\text{Ra Pr}}$.
In practice, we find that our non-rotating simulations equilibrate in $\sim 2.5\sqrt{\text{Ra}\,\text{Pr}}/\text{Nu}$ nondimensional time units (see section \ref{sec:time_evolution_classic_ICs}), and our rotating simulation equilibrates in $\sim (2/3)\sqrt{\text{Ra}\,\text{Pr}}/\text{Nu}$ time units (see section \ref{sec:results_rotating}).

@@

\newpage
\subsection{TT-to-FT \& Nu-based ICs: rapidly equilibrated FT simulations}
\label{sec:tt-to-ft}

\begin{figure}
\includegraphics[width=\textwidth]{./figs/rbc_restart_description.pdf}
\caption{ 
	(Left three panels) Time traces \ea{(in units of $\tau_{\text{ff, ev}}$)} of scalar quantities, which have been averaged over a rolling time window of 25\ea{$\tau_{\text{ff, ev}}$}, are shown for a simulation with Ra$_{\Delta T} = 10^9$ which starts with TT boundary conditions and then is switched to FT boundary conditions \ea{with Ra$_{\partial_z T} = 4.83 \times 10^{10}$}.
	The time of the change of boundary conditions is denoted by the vertical black line.
	(Top panel) The evolution of Ra is shown; Ra$_{\partial_z T}/(4.83\times 10^{10})$ is shown as a dashed-dot line, while Ra$_{\Delta T}/10^9$ is shown as a solid line.
	The mean value of temperature gradient (middle panel) and temperature (bottom panel) at the bottom boundary are also shown.
	In the TT initial state, the temperature is held constant at a value of 1 and the temperature derivative fluctuates around a value of $\text{Nu}$.
	In the FT final state, the temperature derivative is held constant at a value of -1 and the temperature value fluctuates around a value of $\text{Nu}^{-1}$.
	(right four panels) PDFs are shown which compare TT-to-FT dynamics (black PDFs) to dynamics from \ea{the classic-IC} FT case from Fig.~\ref{fig:rbc_scalar_comparisons} (green PDFs).
	We display the temperature field (upper left), enstrophy (upper right), nonlinear convective enthalpy flux (bottom left), and vertical velocity (bottom right).
\label{fig:rbc_restart_description} }
\vspace{-0.25cm}
\end{figure}

@@

\ea{
\subsubsection{Nu-based ICs}
The time evolution of FT simulations with Nu-based ICs is similar to the time evolution of TT simulations with classic ICs, with a few caveats.
The interior temperature gradient ($\grad_{\text{interior}}$ in Eqn.~\ref{eqn:nu_based_gradT}) is slightly positive at high Ra, which means that the interior is marginally thermally stable while the boundary layers are thermally unstable.
During the convective transient, plumes from the boundaries eat away at this interior stratification over a few tens of freefall timescales, after which time the interior is well mixed and a classic roll solution is achieved.
This behavior is quite different from classic ICs, in which the full domain is initially unstable and a roll solution is obtained immediately after the onset of nonlinear convection.
Regardless, the time required for the temperature field to reach statistical equilibrium is a few tens of freefall times rather than a few thousands of freefall times for classic ICs.
Per table \ref{table:pdf_values}, the evolved state of FT simulations with Nu-based ICs and classic ICs are very similar.
}

\subsection{\ea{Discussion of 2D results}} 
We note briefly that \ea{Nu-based and} TT-to-FT \ea{ICs are} only \ea{two} of many ways of accelerating the thermal relaxation of an FT simulation.
We discuss other mechanisms, and explore one in detail, in our previous work \cite{anders&all2018}.
We note however that the TT-to-FT setup described here is likely the least complicated mechanism for achieving rapid relaxation in a simplified RBC setup that we are aware of.
The successful degree with which this mechanism reproduces the evolved dynamics suggests that thermal relaxation occurs in two parts:
\begin{enumerate}
\item Changes to the simulation energy reservoir, and
\item Restratification of the experiment.
\end{enumerate}
The thermal energy reservoir of TT simulations does not change between the initial and final state.
The rapid relaxation of TT simulations\ea{, as well as FT simulations with Nu-based and TT-to-FT ICs}, therefore suggests that experimental restratification occurs rapidly in RBC. 
The long rundown of \ea{classic-IC} FT experiments on display in Fig.~\ref{fig:rbc_scalar_comparisons} is entirely due to the energy reservoir (the temperature jump across the domain) drifting over time.

\ea{
As a final note, we find that measures of the velocity field (e.g., the kinetic energy and Pe) take a few hundred freefall timescales to relax to their final value in TT and FT (Nu-based IC) simulations in 2D at high values of Ra$_{\Delta T}\,\gtrsim 10^{8.67}$.
This velocity field relaxation happens despite instant \emph{thermal} relaxation of the temperature field for these simulations.
We find that the kinetic energy increases by less than a factor of two from its initial post-transient value to its final value in the statistically-stationary state.
}

@@

\ea{
\subsection{3D verification of 2D results}
It would be prohibatively expensive to timestep through the thermal relaxation of a 3D classic-IC FT simulation, even for our least turbulent Ra$_{\Delta T} \approx 10^8$ case.
We anticipate that the transient state would require a spectral coefficient resolution of $\gtrsim 512^3$ for adequate resolution, a factor of 64 times more coefficients than our comparable 2D case at 2048x1024.
This 2D case cost $1.21 \times 10^5$ cpu-hours, so we estimate that comparable 3D case would cost O(10 million) cpu-hours.
We found in previous work \cite{anders&all2018} at lower Ra$_{\partial_z T}$ that 3D classic-IC FT cases exhibited the same thermal rundown as 2D cases.
Here, we will focus on 3D comparisons of FT and TT simulations in the statistically-stationary state through the use of TT-to-FT initial conditions.

We find that equilibrated FT and TT simulations are analogous in a volume-averaged sense: measurements of Nu (31.4 for TT, 31.9 for FT) and Pe ($1.78 \times 10^3$ for both) are nearly indistinguishable between the two cases.
As in 2D, we find no difference between the PDFs of evolved quantities like the nonlinear transport ($wT$) between TT and FT simulations, but we do find differences between the evolved temperature fields ($T$).
In Fig.~\ref{fig:rbc_3D_panels}, we show dynamical slices of $T$ from 3D TT and FT simulations.
Near the top boundary and in the interior, the TT and FT dynamics are quite similar.
However, near the bottom boundary, the warm upflows in FT simulations are hotter relative to their counterparts in TT simulations.

@@

\ea{
Rotating convection is an excellent testbed for our TT-to-FT method.
In rotating convection, as Ra increases at a fixed value of the Ekman number (Ek), flows transition from the rotationally constrained to the rotationally unconstrained regime.
The scaling of Nu~vs.~Ra changes drastically between these two regimes, and is some blend of the two in the intermediary, marginally constrained regime.
Furthermore, the precise scaling law attained in the rotationally constrained regime differs as Ek changes, and these scaling laws are less straightforward and well-understood than their non-rotationally-constrained counterparts \cite{king&all2009, schmitz&tilgner2009, zhong&ahlers2010, julien&all2012, stevens&all2013, ecke2015, grooms2015, grooms&whitehead2015,  plumley&julien2019}.
As a result, the powerlaw used for Nu-based ICs would have to be a complex function of Ek, Ra, and boundary conditions for rotating RBC.
However, TT-to-FT should work generally for all parameters, so long as a TT simulation can be performed.
Here we explore the parameter space that the classic-IC FT simulation explores as it thermally relaxes and demonstrate that the TT-to-FT approach leads to rapid convergence for a rotating simulation.
}

We now study 3D rotating RBC with Ek$ \,= 10^{-6}$.
These simulations employ stress free boundary conditions which allow for the generation of mean flows such as large scale vortices (LSV) \ea{\citep{stellmach&all2014, rubio&all2014, guervilly&all2014, guervilly&hughes2017, favier&all2014, favier&all2019, couston&all2019}}.
We study a TT case at Ra$_{\Delta T} = 2.75\times 10^9$, and a classic-IC FT case at $\text{Ra}_{\partial_z T} = 2.1 \times 10^{10}$ (the supercriticality of the TT case is $\sim 3$).
\ea{We then take the TT case, do a TT-to-FT simulation, and compare the results of the TT-to-FT simulation to the classic-IC FT case.}

In the left three panels of Fig.~\ref{fig:rotating_panels}, we compare the time evolution of \ea{a classic-IC} FT and TT case.
The top left panel shows the evolution of Ra$_{\partial_z T}$ and Ra$_{\Delta T}$.
Even in the presence of strong rotation, the TT simulation immediately equilibrates, but the FT case takes thousands of freefall times to achieve thermal relaxation.
In the middle panel, we show the evolution of Ro; the evolved flows in both simulations exhibit rotationally constrained dynamics with Ro $\,\approx 0.1$, but the flows in the FT simulation relax to this state from an initially unconstrained state (Ro $\,\approx 1$).
This implies that the thermal relaxation process can walk through the parameter space of flow balances (e.g., the balance between Inertial and Coriolis forces) in addition to the Ra$_{\Delta T}$ parameter space.
In the bottom panel, we display the evolution of Pe over time.
Strangely, the peak value of Pe occurs a few hundred freefall times after the convective transient.
After achieving this peak value, Pe monotonically decreases toward its relaxed state.
\ea{We find that the thermal relaxation of this case takes $\sim (2/3)\sqrt{\text{Ra}_{\partial_z T}\text{ Pr}}/\text{Nu}$ nondimensional freefall time units.}

\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/rotating_panels.pdf}
\caption{ 
	(Left three panels) Time traces of scalar quantities in a \ea{classic-IC} FT (orange) and TT (purple) simulation are shown.
	All traces have been averaged over a rolling window of 50 freefall time units to increase the clarity of the evolutionary trend.
	(Top panel) Ra, normalized by the input Ra$_{\Delta T}$ = 2.75$\,\times 10^9$ of the TT simulation.
	(Middle panel) Ro evolution of both simulations; the bulk flow of the FT simulation transitions from a marginally rotationally unconstrained state to a constrained state, while the TT simulation is always rotationally constrained.
	(Bottom panel) Pe evolution of the simulations is shown, normalized by the mean value measured over the last 500 freefall times of the TT simulation.
	(Upper right panel) Parameter space of Nu vs.~Ra$_{\Delta T}$ in rotating convection.
	Circular and diamond data points are respectively simulations and experimental data points from ref.~\cite{cheng&all2015}.
	The color of the data points signifies the Ekman number of the points, and grey diamonds are non-rotating.
	Data from our Ek$\,=\,10^{-6}$ FT experiment are shown as a thick orange line with a cyan interior, where the black arrows give the direction of time.
	Some TT simulations are shown as purple circles with a cyan interior, and the TT case which corresponds to the relaxed state of the FT simulation is shown as a star.
	(Bottom right panels) Snapshots of the vertically integrated z-component of the vorticity from the FT simulation.
	At early times (left panel), a powerful large scale vortex with positive vorticity develops.
	This vortex slowly decays and becomes a vortex pair (middle panel), as seen in ref.~\cite{stellmach&all2014}.
	In the converged state, we see oscillatory behavior between this vortex pair behavior and jets (right panel).
	The TT case exhibits the oscillatory behavior between vortex pairs and jets throughout its whole evolution.
	The three vertical black lines in the left panels signify the times at which these snapshots are taken.
\label{fig:rotating_panels} }
\end{figure}

@@

\ea{
We performed a TT-to-FT simulation starting from the evolved state of the TT simulation, and its time evolution matched that seen in Fig.~\ref{fig:rbc_restart_description}.
There was no long thermal evolutionary timescale.
In the equilibrated classic FT simulation, we measured Nu$=7.90 \pm 0.01$, Pe$=1.70 \times 10^3$, and Ro$=0.115 \pm 0.006$.
In the TT-to-FT simulation, we measured Nu$=7.83 \pm 0.01$, Pe$=1.65 \times 10^3$, and Ro=$0.114 \pm 0.005$.
These Ro measurements are indistinguishable, Nu measurements are within 1\% of each other, and Pe measurements are within 3\% of each other, representing excellent agreement.

The difference in computational cost between classic-IC FT simulations and the TT/TT-to-FT simulations are even more striking here than in our previous examples.
}
The TT simulation shown in the left panels of Fig.~\ref{fig:rotating_panels} only cost $2.2 \times 10^3$ cpu-hours to run 5000 freefall times.
\ea{The TT-to-FT simulation cost an additional $1.28 \times 10^3$ cpu-hours to run 1000$\tau_{\text{ff, ev}}$.}
By comparison, the cost of the FT simulation shown in the same panels was \emph{roughly three orders of magnitude larger}---$2.3 \times 10^6$ cpu-hours.
The TT \ea{and TT-to-FT} simulations \ea{had a} coefficient resolution \ea{of} $128^3$.
The FT simulation's initial resolution required to resolve the convective transient was $512\times384^2$ coefficients.
We reduced the resolution to $256\times384^2$ after 100 freefall times, and then later to $128\times384^2$ after $\sim3.3 \times 10^3$ freefall times.
At each of these times, we found that lowering the \emph{horizontal} coefficient resolution of the simulation did not reproduce the simulation solution with fidelity.
This suggests that small scale turbulent velocity structures---which are injected by the vigorous transient and perhaps associated with the LSV---are long lived throughout the thermal evolution of the simulation.
